### 2014-08-18 analysis/correlation_test.Rmd

Context Permeability Convergence Correlation
============================================
In this report we will do a quick analysis of the correlation between simulation convergence and 
network properties. We will compute the **pearman's rank correlation coefficient** to analyse our data.

```{r}
source("../R/experiment_utils.R")
library(reshape)
library(reshape2)
library(data.table)

# Correlations with significance levels
library(Hmisc)

```

## K-Regular Networks ##
```{r}
#read the network properties
props <- read.table(file="../data/context_permeability/regular_networks/network_properties.csv", sep = ",", header = T)
head(props)

apl_data <- melt(props, id=c("k","networks","run"),measure.vars=c("average.path.length"))
clustering_data <- melt(props, id=c("k","networks","run"),measure.vars=c("clustering.coefficient"))

average_apl_data <- dcast(apl_data, k + networks ~ variable, mean)
average_clustering_data <- dcast(clustering_data, k + networks ~ variable, mean)
head(average_apl_data)
head(average_clustering_data)

#read the convergence data
params <- read.exp.parameters(param_file_name="../data/context_permeability/regular_networks/convergence/param-space_2014-08-01_00:09:09experiment:0_.csv")

#read the adata related to consensus was achieved before

convergence_data <- fread("../data/context_permeability/regular_networks/convergence/consensus_achieved.csv")
convergence_data <- as.data.frame(convergence_data)[,-ncol(convergence_data)]
convergence_data <- merge(convergence_data, params, by="cfg.id")

convergence_data <- melt(convergence_data,id.vars=c("cfg.id","run","num.networks","network.0.k"), measure.vars=c("consensus-achieved"))

convergence_data <- aggregate(data=convergence_data, convergence_data$value~convergence_data$num.networks + convergence_data$network.0.k, FUN=function(x) sum(x)/length(x)*100 )

colnames(convergence_data) <- c("networks", "k", "convergence")
head(convergence_data)

table <- cast(convergence_data, networks~k, value="convergence")
latex_table <- tabular.cast_df(table)

full_data <- merge(convergence_data,average_apl_data, by=c("networks","k"))
full_data <- merge(full_data, average_clustering_data, by=c("networks","k"))
head(full_data)
```

```{r}
pairs(full_data)

#Spearman's correlation applies to ranks and so provides a measure of a monotonic relationship between two continuous random variables. It is also useful with ordinal data and is robust to outliers (unlike Pearson's correlation).
apl_test <- cor.test(normal.range(full_data$convergence), normal.range(full_data$average.path.length), type="spearman", conf.level = 0.95)
library(ggplot2)
clustering_test <- cor.test(normal.range(full_data$convergence), normal.range(full_data$clustering.coefficient), type="spearman", conf.level = 0.95)
qplot(normal.range(full_data$convergence), normal.range(full_data$average.path.length)) +xlab("Convergence Ratio") + ylab("Normalised Average Path Length")

apl_test
clustering_test
```

## Scale-free Networks ##
```{r}
#read the network properties
props <- read.table(file="../data/context_permeability/scale-free_networks/network_properties.csv", sep = ",", header = T)
head(props)

apl_data <- melt(props, id=c("d","networks","run"),measure.vars=c("average.path.length"))
clustering_data <- melt(props, id=c("d","networks","run"),measure.vars=c("clustering.coefficient"))

average_apl_data <- dcast(apl_data, d + networks ~ variable, mean)
average_clustering_data <- dcast(clustering_data, d + networks ~ variable, mean)
head(average_apl_data)
head(average_clustering_data)

#read the convergence data
params <- read.exp.parameters(param_file_name="../data/context_permeability/scale-free_networks/convergence/param-space_2014-08-01_00:47:12experiment:0_.csv")

#read the adata related to consensus was achieved before

convergence_data <- fread("../data/context_permeability/scale-free_networks/convergence/consensus_achieved.csv")
convergence_data <- as.data.frame(convergence_data)[,-ncol(convergence_data)]
convergence_data <- merge(convergence_data, params, by="cfg.id")

convergence_data <- melt(convergence_data,id.vars=c("cfg.id","run","num.networks","network.0.d"), measure.vars=c("consensus-achieved"))
convergence_data <- aggregate(data=convergence_data, convergence_data$value~convergence_data$num.networks + convergence_data$network.0.d, FUN=function(x) sum(x)/length(x)*100)
colnames(convergence_data) <- c("networks", "d", "convergence")
head(convergence_data)

full_data <- merge(convergence_data,average_apl_data, by=c("networks","d"))
full_data <- merge(full_data, average_clustering_data, by=c("networks","d"))
head(full_data)
```

```{r}
pairs(full_data)

#Spearman's correlation applies to ranks and so provides a measure of a monotonic relationship between two continuous random variables. It is also useful with ordinal data and is robust to outliers (unlike Pearson's correlation).
apl_test <- cor.test(normal.range(full_data$convergence), normal.range(full_data$average.path.length), type="spearman",conf.level = 0.95)
library(ggplot2)
clustering_test <- cor.test(normal.range(full_data$convergence), normal.range(full_data$clustering.coefficient), type="spearman",conf.level = 0.95)
qplot(normal.range(full_data$convergence), normal.range(full_data$average.path.length)) +xlab("Convergence Ratio") + ylab("Normalised Average Path Length")

apl_test
clustering_test
```